Long range correlations and their event scale dependence in high multiplicity pp collisions at $\sqrt{s} = 13$~TeV

### 2020-06-27 sec_intro.tex
Updates from Overleaf

Strong collectivity observed in the azimuthal correlations of particles emitted over a wide pseudorapidity range, in high-energy nucleus-nucleus collisions at RHIC~\cite{Adams:2005dq,Adcox:2004mh,Arsene:2004fa,Back:2004je} and LHC~\cite{Abelev:2012di, Abelev:2014pua, ATLAS:2011ah}, indicates the formation of a strongly interacting matter called the quark-gluon plasma (QGP), which exhibits hydrodynamic behavior (see the reviews~\cite{Romatschke:2007mq,Jeon:2015dfa,Romatschke:2017ejr}). The recent efforts are focused on constraining the transport properties of the QGP in hydrodynamic models~\cite{Niemi:2015qia,Bernhard:2016tnd,Bernhard2019} along with few advanced experimental efforts~\cite{ALICE:2016kpq,Acharya:2017gsw,Acharya:2017zfg,Acharya:2020taj}.
In recent years similar collective behaviors via long-range azimuthal correlations are also observed for small systems with high final-state particle multiplicity such as proton-proton (pp)~\cite{Aad:2015gqa,Khachatryan:2015lva,Khachatryan:2016txc,Acharya:2019vdf}, proton-nucleus (pA)~\cite{Abelev:2012ola,Aad:2014lta,Aaboud:2016yar,Khachatryan:2016ibd}, and lighter nucleus-nucleus systems~\cite{PHENIX:2018lia,Aidala:2017ajz}, revealing strong indications for collective flow with hydrodynamic characteristics even in small systems, though the volume and lifetime of the medium produced by such a system are expected to be small.

>>>>

Strong collectivity observed in the azimuthal correlations of particles emitted over a wide pseudorapidity range, in high-energy nucleus-nucleus collisions at RHIC~\cite{Adams:2005dq,Adcox:2004mh,Arsene:2004fa,Back:2004je} and LHC~\cite{Abelev:2012di, Abelev:2014pua, ATLAS:2011ah}, indicates the formation of a strongly interacting matter called the quark-gluon plasma (QGP), which exhibits hydrodynamic behavior (see the reviews~\cite{Romatschke:2007mq,Jeon:2015dfa,Romatschke:2017ejr}). The recent efforts are focused on constraining the transport properties of the QGP in hydrodynamic models~\cite{Niemi:2015qia,Bernhard:2016tnd,Bernhard2019} along with few advanced experimental efforts~\cite{ALICE:2016kpq,Acharya:2017gsw,Acharya:2017zfg,Acharya:2020taj}.
In recent years similar collective behaviors via long-range azimuthal correlations are also observed for small systems with high final-state particle multiplicity such as proton-proton (pp)~\cite{Aad:2015gqa,Khachatryan:2015lva,Khachatryan:2016txc,Acharya:2019vdf}, proton-nucleus (pA)~\cite{Abelev:2012ola,Aad:2014lta,Aaboud:2016yar,Khachatryan:2016ibd}, and lighter nucleus-nucleus systems~\cite{PHENIX:2018lia,Aidala:2017ajz}, revealing strong indications for collective flow with hydrodynamic characteristics even in small systems, though the volume and lifetime of the medium produced by such a system are expected to be small~\cite{Busza:2018rrf,Nagle:2018nvi}.
